\vspace{-.2in} Squint: Lossy Hierarchical Compression on Symbol Streams \vspace{-.1in}

### 2012-12-12 tex/paper.tex
Subsubsection format
Finishing Implementation section

\sectionfont{\large}
\subsectionfont{\normalsize}
\subsubsectionfont{\normalfont}

@@

We chose to implement a list library rather than use Ruby's native arrays in
order to achieve better runtime for the splice operations.  Both \Sequitur and
our implementation of Kieffer and Yang's reductions require us to splice out a
subsequence of terminals from a rule, or to insert a subsequence (e.g., as a
result of inlining a singleton rule) into a rule.

Grammar Data Structure
Linked Lists vs Arrays, Caching Expansions, Native Levenshtein

@@

Perhaps the most crucial optimization we implement deals with the recursive
expansion of nonterminals into terminal strings.

@@

What is their purpose? Why do we want them? What are they?  Grammar reductions
are a way of taking a CFG, generated possibly through ad-hoc techniques, and
systematically removing internal redundancies like duplicate bigrams in
\Sequitur. They make the CFG representation smaller but don't introduce any
loss. We want them since our lossifier algorithms often make our grammars
redundant, even though the each step we take in the lossifier makes the
grammars smaller.

>>>>

\sectionfont{\large}
\subsectionfont{\normalsize}
\subsubsectionfont{\normalfont\itshape}

@@

We chose to implement a list library rather than use Ruby's native arrays in
order to achieve better runtime for the splice operation.  Both \Sequitur and
our implementation of Kieffer and Yang's reductions require us to splice out a
subsequence of terminals from a rule, or to insert a subsequence (e.g., as a
result of inlining a singleton rule) into a rule.  Moreover, our initial
implementation of the \texttt{replace} operation---which replaces all
occurrences of a nonterminal with another---initially employed rule inlining to
prevent the creation of cycles.  Our first iteration implementation used native
arrays, and we found that linked lists did result in an empirical improvement
in runtime.

@@

Perhaps the most crucial optimization we implement deals with the recursive
expansion of nonterminals into terminal strings---the \texttt{expand}
operation.  An extraordinarily na\"ive implementation, in which we simply
recurse into every rule we find, could have runtime as bad as $O(\prod_{R \in
G} |R|)$, the product of the lengths of all rules in the grammar $G$.  We can
easily reduce this to $O(\sum_{R \in G} |R|)$ by caching the expansion of each
nonterminal when we complete it, and using the cached result for later
instances.

We further leverage this caching optimization, however, by preserving the
expansion cache across calls to \texttt{expand}.  This optimization is critical
to achieving acceptable runtime for the \Similarity and \Cluster algorithms
because of their heavy reliance on expanding.  Each iteration of the algorithm
requires expanding all nonterminals.  Since \texttt{replac}ing a nonterminal A
with B changes the expansion of all rules containing A, we clear the expand
cache whenever replace is called.  This is the key reason for \Cluster's
superior runtime relative to \Similarity---\Similarity needs to expand all
nonterms for each individual \texttt{replace} operation, whereas \Cluster
clusters \texttt{replace} operations, thereby reducing the number of calls and
effective overhead of \texttt{expand}.

In addition to this optimization, we also use a native C implementation of the
Levenshtein function.  We do not compare this to any Ruby implementation, but
we suspect it has some nontrivial impact on runtime due to the volume of calls
to the function.

@@

Kieffer and Yang's grammar reductions are a collection of rules which, when
applied iteratively to a grammar, systematically eliminate internal
redundancies, such as duplicate symbol subsequences.  They often reduce the
size of CFGs by factoring out repeated patterns as new rules and by eliminating
unnecessary rules, but they introduce no loss.

Since our lossification techniques make no guarantees about the form of the
resultant grammar, unlike, say, \Sequitur, we produced a na\"ive implementation
of these reduction rules in order to make our lossy grammars as small as
possible.  As we shall see in the next section, this simple reduction engine
has poor runtime and does not scale very well.
